---
title: "Homework 5 Iteration"
author: "Minghui Wang" 
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---
This assignment reinforces ideas in [Iteration](https://p8105.com/topic_iteration.html).

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
set.seed(123)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1
Write a function that, for a fixed group size, randomly draws “birthdays” for each person; checks whether there are duplicate birthdays in the group; and returns TRUE or FALSE based on the result.
```{r}
bday_sim = function(n) {

  bdays = sample(1:365, size = n, replace = TRUE)
  
  duplicate = length(unique(bdays)) < n

  return(duplicate)
}
```

Run this function 10000 times for each group size between 2 and 50. For each group size, compute the probability that at least two people in the group will share a birthday by averaging across the 10000 simulation runs.
```{r}
sim_res = 
  expand_grid(
    n = 2:50,
    iter = 1:10000
  ) |> 
  mutate(res = map_lgl(n, bday_sim)) |> 
  group_by(n) |> 
  summarize(prob = mean(res))
```
 Make a plot showing the probability as a function of group size, and comment on the results.
```{r}
sim_res |> 
  ggplot(aes(x = n, y = prob )) + 
  geom_line()+
  labs(
    x= "Group Size (n)",
    y= "Probability",
    title = "Probability of Duplicate Birthdays as Group Size Increases"
  )
```

**Comment** : The plot shows an increasing trend in the probability that at least two people share a birthday as the group size grows. The probability starts near zero for small groups but rises quickly, reaching around 50% by a group size of 23, and approaching nearly 100% by a group size of 50. The increasing rate is biggest around a group size of 23 and smallest around 0 and 50. This plot demonstrates how the likelihood of shared birthdays increases as more people are added.

# Problem 2
Define the function to save 𝜇̂ and the p-value arising from a test of 𝐻:𝜇=0 using 𝛼=0.05.
```{r}
# Set parameters
n<- 30
sigma <- 5
alpha <- 0.05
```

```{r}
# Define t_test function  
t_test = function(mu) {
  
  sample_data = rnorm(n, mean = mu, sd = sigma)
  
  t_stat = t.test(sample_data, mu = 0) |>
    broom::tidy() |>
    select(estimate, p.value)
  
  return(t_stat)
}
```

```{r}
# Simulation for 5000 datasets for each mu value and store results
sim_result_df =
  expand_grid(
    mu = 0:6,
    iter = 1:5000 )|>
  mutate(
    estimate_df = map(mu, t_test)
  )|> 
  unnest(estimate_df)
```



# Problem 3
```{r}
```